**The status of the microbial census: an update**

### 2015-11-23 doc/notebook/Bacterial_data_acquisition.Rmd
Change metadat files to be *.v123.metadata
Fix path issue btwn pre.cluster and cluster.split

* ARB_color != 1 & tax_rdp == "*Bacteria*" (N=1,384,528)
* ARB_color != 1 & tax_greengenes == "*Bacteria*" (N=1,123,279)
* ARB_color != 1 & tax_slv == "*Bacteria*" (N=1,526,523)
* ARB_color != 1 & tax_embl == "*Bacteria*" (N=1,335,715)
* ARB_color != 1 & (tax_rdp | tax_greengenes | tax_embl | tax_slv) (N=1,529,365)

@@

* Return to Search and Query window
* Change search field to "ARB_color", enter "1", and click the equal sign to be not-equal
* Change the second search field to "tax_slv" and set it to "*Bacteria*"
* Click "Mark Listed, Unmark Rest"
* Change the first search field to "tax_slv" and set it to "*Chloroplast*", set to "not equal"
* Change the second search field to "tax_slv" and set it to "*Mitochondria*", set to "not equal"
* Set the two Query operators to "and"
* Click "Keep species" and "that match query"
* Hit "Search"
* Hit Mark Listed Unmark Rest (N=1,515,024)
* Click "Write to Fields of Listed", select the "remark" field, and enter "good bacteria"
* Click "Write"
* In main ARB window go File->Export->Export to external format
* Select Compress -> "Vertical Gaps" and fasta_mothur.eft as the format
* Rename `noname.fasta` to `data/mothur/bacteria.fasta`

@@

* Go Tree -> NDS
* Click "name", "acc", "tax_slv". The "tax_slv" field should have 250 characters
* Unclick everything else
* Click "Close"
* Go File->Export->Export fields
* Set the file name to "data/mothur/bacteria.taxonomy" and Column output to "TAB separated"
* Click "SAVE"

@@

```
            Start	End		NBases	Ambigs	Polymer	NumSeqs
Minimum:    1		10725	900		0		4	1
2.5%-tile:	1003	14692	1274	0		5	37876
25%-tile:	1011	14865	1355	0		5	378757
Median: 	1044	15082	1387	0		5	757513
75%-tile:	1048	15672	1455	0		6	1136269
97.5%-tile:	1815	15771	1539	2		7	1477149
Maximum:	7036	17738	3469	64		38	1515024
Mean:		1104.14	15254.4	1404	0.181	5.509
# of Seqs:	1515024
```

@@

```{r find_screening_points}
data <- read.table(file="data/mothur/bacteria.summary", header=T, row.names=1)
quantile(data$start, probs=seq(0.9,1,0.01))
start <- 2000

quantile(data$end, probs=seq(0,0.1,0.01))
end <- 14853

@@

trimmed <- data[data$start <= start & data$end >= end &
                    data$nbases > min_length & data$ambigs <= max_ambig,]
nrow(trimmed)
summary(trimmed$nbases)
```

@@

```{r screen_data, engine="bash"}
mothur "#screen.seqs(fasta=data/mothur/bacteria.fasta, taxonomy=data/mothur/bacteria.tax, start=2000, end=14853, maxambig=2, minlength=1200, outputdir=data/mothur/, processors=8);"
```

@@

```{r single_cell, engine="bash"}
#wget -N -P data/references/ http://www.mothur.org/w/images/2/24/Trainset10_082014.pds.tgz
#tar xvzf data/references/Trainset10_082014.pds.tgz -C data/references/
#mv data/references/trainset10_082014.pds/trainset10_082014.* data/references/
#rm -rf data/references/trainset10_082014.pds

mothur "#set.dir(output=data/mothur);
        align.seqs(fasta=data/mothur/single_cell.bacteria.fasta, reference=data/mothur/bacteria.good.fasta);
        set.dir(input=data/mothur);
        screen.seqs(fasta=current, start=2000, end=13000, maxambig=2, minlength=1200);
        classify.seqs(fasta=current, reference=data/mothur/bacteria.good.fasta, taxonomy=data/mothur/bacteria.good.tax, cutoff=80)"
```

@@

#need to make the taxonomy file match the fasta and names files...
cut -f 1 data/mothur/all_bacteria.filter.unique.precluster.names > precluster.accnos
mothur "#set.dir(input=data/mothur, output=data/mothur);
        get.seqs(taxonomy=all_bacteria.taxonomy, accnos=precluster.accnos)"
```

@@

#correct typos:
metadata$category[metadata$category == 'AF'] <- 'AE'
metadata$category[metadata$category == 'AQBF'] <- 'AQB'
metadata$category[metadata$category == 'AQD'] <- 'AQO'
metadata$category[metadata$category == 'AQHS'] <- 'AQH'
metadata$category[metadata$category == 'AQS'] <- 'AQMS'
metadata$category[metadata$category == 'AWF'] <- 'AQF'
metadata$category[metadata$category == 'DS'] <- 'BS'
metadata$category[metadata$category == 'S'] <- 'SO'
metadata$category[metadata$category == 'SI'] <- 'BI'
metadata$category[metadata$category == 'V'] <- 'BO'
metadata$category[metadata$category == 'VD'] <- 'BD'
metadata$category[metadata$category == 'ZOZO'] <- 'ZO'
metadata$category[metadata$category == 'AH'] <- 'AQH'
metadata$category[metadata$category == 'AQOS'] <- 'AQS'
metadata$category[metadata$category == 'BS'] <- 'BD'
metadata$category[metadata$category == 'AQS'] <- 'AQFS'
metadata$category[metadata$category == 'ZL'] <- 'ZV'
metadata$category[metadata$category == 'SR'] <- 'PR'
metadata$category[metadata$category == 'M'] <- 'AQM'
metadata$category[metadata$category == 'AQFM'] <- 'AQFS'
metadata$category[metadata$category == 'SV'] <- 'ZV'

#... and we output the metadata file...
write.table(metadata, file="data/mothur/bacteria.final.metadata", sep="\t")

>>>>

* ARB_color != 1 & tax_rdp == "Bacteria*" (N=1,376,114)
* ARB_color != 1 & tax_greengenes == "k__Bacteria*" (N=1,119,394)
* ARB_color != 1 & tax_slv == "Bacteria*" (N=1,516,624)
* ARB_color != 1 & tax_embl == "Bacteria*" (N=1,331,557)

@@

* Return to Search and Query window
* Change search field to "ARB_color", enter "1", and click the equal sign to be not-equal
* Change the second search field to "tax_slv" and set it to "Bacteria*"
* Click "Mark Listed, Unmark Rest"
* Change the first search field to "tax_slv" and set it to "*Chloroplast*", set to "not equal"
* Change the second search field to "tax_slv" and set it to "*Mitochondria*", set to "not equal"
* Set the two Query operators to "and"
* Click "Keep species" and "that match query"
* Hit "Search"
* Hit Mark Listed Unmark Rest (N=1,505,125)
* In main ARB window go File->Export->Export to external format
* Select Compress -> "Vertical Gaps" and fasta_mothur.eft as the format
* Rename `noname.fasta` to `data/mothur/bacteria.fasta`

@@

* Go Tree -> NDS
* Click "name", "acc", "tax_slv". The "tax_slv" field should have 250 characters
* Unclick everything else
* Click "Close"
* Go File->Export->Export fields
* Set the file name to "data/mothur/bacteria.taxonomy.nds" and Column output to "TAB separated"
* Click "SAVE"

@@

```
				Start   End     NBases  Ambigs  Polymer NumSeqs
Minimum:        1       12530   1200    0       4       1
2.5%-tile:      1003    14538   1282    0       5       37629
25%-tile:       1011    14707   1356    0       5       376282
Median:         1044    14926   1387    0       5       752563
75%-tile:       1048    15514   1455    0       6       1128844
97.5%-tile:     1795    15613   1539    2       7       1467497
Maximum:        5300    17580   3469    64      38      1505125
Mean:   		1099.25 15100.4 1405.16 0.180003        5.50868
# of Seqs:      1505125
```

@@

```{r find_screening_points}
data <- read.table(file="data/mothur/bacteria.summary", header=T, row.names=1)

quantile(data$start, probs=seq(0.9,1,0.01))
start <- 2000

quantile(data$end, probs=seq(0,0.1,0.01))
end <- 14459

@@

trimmed <- data[data$start <= start & data$end >= end &
                    data$nbases > min_length & data$ambigs <= max_ambig,]

nrow(trimmed) #[1] 1431222

summary(trimmed$nbases)
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
#1201    1359    1390    1409    1457    3469
```

@@

```{r screen_data, engine="bash"}
mothur "#screen.seqs(fasta=data/mothur/bacteria.fasta, taxonomy=data/mothur/bacteria.tax, start=2000, end=14459, maxambig=2, minlength=1200, outputdir=data/mothur/, processors=8);"
```

@@

```{r single_cell, engine="bash"}
mothur "#set.dir(output=data/mothur);
        align.seqs(fasta=data/mothur/single_cell.bacteria.fasta, reference=data/mothur/bacteria.good.fasta);
        set.dir(input=data/mothur);
        screen.seqs(fasta=current, start=2000, end=13000, maxambig=2, minlength=1200);
        classify.seqs(fasta=current, reference=data/mothur/bacteria.good.fasta, taxonomy=data/mothur/bacteria.good.tax, cutoff=80)"
```

@@

#need to make the taxonomy file match the fasta and names files...
cut -f 1 data/mothur/all_bacteria.filter.unique.precluster.names > data/mothur/precluster.accnos
mothur "#set.dir(input=data/mothur, output=data/mothur);
        get.seqs(taxonomy=all_bacteria.taxonomy, accnos=precluster.accnos)"
```

@@

#correct typos:
metadata$category[metadata$category == 'AF'] <- 'AE'
metadata$category[metadata$category == 'AQBF'] <- 'AQB'
metadata$category[metadata$category == 'AQD'] <- 'AQO'
metadata$category[metadata$category == 'AQHS'] <- 'AQH'
metadata$category[metadata$category == 'AQS'] <- 'AQMS'
metadata$category[metadata$category == 'AWF'] <- 'AQF'
metadata$category[metadata$category == 'DS'] <- 'BS'
metadata$category[metadata$category == 'S'] <- 'SO'
metadata$category[metadata$category == 'SI'] <- 'BI'
metadata$category[metadata$category == 'V'] <- 'BO'
metadata$category[metadata$category == 'VD'] <- 'BD'
metadata$category[metadata$category == 'ZOZO'] <- 'ZO'
metadata$category[metadata$category == 'AH'] <- 'AQH'
metadata$category[metadata$category == 'AQOS'] <- 'AQS'
metadata$category[metadata$category == 'BS'] <- 'BD'
metadata$category[metadata$category == 'AQS'] <- 'AQFS'
metadata$category[metadata$category == 'ZL'] <- 'ZV'
metadata$category[metadata$category == 'SR'] <- 'PR'
metadata$category[metadata$category == 'M'] <- 'AQM'
metadata$category[metadata$category == 'AQFM'] <- 'AQFS'
metadata$category[metadata$category == 'SV'] <- 'ZV'
metadata$category[metadata$category == 'ZI'] <- 'ZA'

#... and we output the metadata file...
write.table(metadata, file="data/process/bacteria.v123.metadata", sep="\t")
